### 2018-04-11 inst/doc/DiceEnterprise.Rmd
Added vignettes

---
title: "Dice Enterprise"
author: "Giulio Morina"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Dice Enterprise package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7, fig.height=4,
  cache=TRUE
)
require("DiceEnterprise")
set.seed(10, "L'Ecuyer-CMRG") 
```

## Introduction

The `DiceEnterprise` package provides an automatic way to construct a Bernoulli Factory/Dice Enterprise algorithm for rational functions between simplices $f: \Delta^m \rightarrow \Delta^v$. That is, given an $m$-sided die, a $v$ sided-one is produced where the probability of each of the $v$ faces is a rational function. More formally, $f(\boldsymbol{p})$ is of the following form:
$$
f(\boldsymbol{p}) = f(p_1,\ldots,p_m) = \frac{1}{C(\boldsymbol{p})}\left(G_1(\boldsymbol{p},\ldots,G_v(\boldsymbol{p})\right)
$$
where each $G_i(\boldsymbol{p})$ and $C(\boldsymbol{p})$ are polynomials with real coefficients.

The Bernoulli Factory is a special case of this setting. In particular, assume a $p$-coin is given. Then, the package can construct an $f(p)$-coin where $f(p)$ is a rational function.

## Bernoulli Factory

The user can define a function that tosses a $p$-coin and returns $1$ with probability $p$ and 2 with probability $1-p$ (notice that **it should not return 1 or 0**). The function should have as input the number `n` of required tosses.  For instance the following function tosses a $\frac{3}{4}$-coin
```{r toss_coin}
toss_coin <- function(n) {
  sample(1:2, size = n, replace = TRUE, prob = c(3/4,1/4))
}
```

Then, the user needs to define the rational function $f(p)$ in a specific format. For instance, assume that 
$$
f(p) = \frac{\sqrt{2}p^3}{(\sqrt{2}-5)p^3+11p^2-9p+3}
$$
so that 
$$
1-f(p) = \frac{-5p^3+11p^2-9p+3}{(\sqrt{2}-5)p^3+11p^2-9p+3}
$$
Then, the user needs to specify the numerators of $f(p)$ and $1-f(p)$ in the following format:
```{r bf_poly}
f_1 <- list(coeff = c(sqrt(2)), power = c(3)) #f(p)
f_2 <- list(coeff = c(-5,11,-9,3), power = c(3,2,1,0)) #1-f(p)
```
Notice that the polynomials are represented as a list: the first element is a vector of coefficients and the second element is a vector describing the powers of $p$. Finally, the user can construct a new Bernoulli Factory running the following code:
```{r bf_construction}
bf <- BernoulliFactory$new(f_1 = f_1, f_2 = f_2) #f_1 = f(p), f_2 = 1-f(p)
```
and toss an $f(p)$-coin by running
```{r bf_toss}
fp_tosses <- bf$sample(n = 10, roll.fun = toss_coin) #Produces 10 tosses of the f(p)-coin
print(fp_tosses)
```

### Diagnosis and extra options

Printing the `BernoulliFactory` object gives information on the dimension of the generated fine and connected ladder. When a `BernoulliFactory` object is created, if the option `verbose` is set equal to `TRUE`, then more information are printed on the different steps of the construction.

```{r bf_diagnosis}
bf <- BernoulliFactory$new(f_1 = f_1, f_2 = f_2, verbose = TRUE)
print(bf)
```
The method `evaluate` allows to evaluate the value of $f(p)$ for a given $p$. 
```{r bf_evaluate}
print(bf$evaluate(3/4))
```

When a sample is requested, Coupling From the Past is used. Notice that in the Bernoulli Factory case (or more in general when the given die has only 2 faces), a monotonic implementation is used. The method `sample` allows to set the option `verbose = TRUE`. In this case the output of `sample` is a list where the first argument is the obtained sample and the second argument is a vector containing the number of rolls required. The same method allows to specify the number of cores used (default 1). Notice that multicore is not supported on Windows. The method `sample` takes also `double_time` as an input. If `double_time = FALSE` (default), the time step of CFTP is increased by one at each iteration leading, otherwise it is doubled.
```{r bf_toss_2}
fp_tosses <- bf$sample(n = 1000, roll.fun = toss_coin, num_cores = 2, verbose = TRUE, double_time= TRUE) #Produces 1000 tosses of the f(p)-coin, using 2 cores and doubling the time step at each iteration of CFTP.
print(table(fp_tosses[[1]])/1000) #Empirical probabilities. Notice that the theoretical ones are given by print(bf$evaluate(3/4))
print(paste0("Average number of tosses required: ",mean(fp_tosses[[2]])))
```
The function `plot.confidence.interval` allows to print the estimated probabilities with 95% confidence interval. The stars indicate the true theoretical values.
```{r bf_plot}
plot.confidence.interval(fp_tosses[[1]],print(bf$evaluate(3/4)))
```

Notice that the package supports only functions defined from $(0,1)$ to $[0,1]$ and not from subset $S \subset (0,1)$. For instance, if the user defines a Bernoulli Factory for the function $f(p) = 2p$, an error is returned:
```{r bf_error, error=TRUE}
bf_amp <- BernoulliFactory$new(f_1 = list(2,1), f_2 = list(c(1,-2),c(0,1)))
```

### Example: Logistic Bernoulli Factory

Consider the function as in Huber
$$
f(p) = \frac{Cp}{1+Cp}
$$
Huber proposes a very clever algorithm that makes use of a thinned Poisson point process and requires on average $C/(1+Cp)$ tosses. If my calculations are correct, the Dice Enterprise can construct a Bernoulli Factory that albeit being always worse, requires at most one extra toss on average. Let's check this:

```{r bf_logistic}
toss.coin.logistic <- function(n,p) {
  sample(1:2, size = n, replace = TRUE, prob = c(p,1-p))
}
C <- 10
p <- 0.01
size_sample <- 10000

bf_logistic <- BernoulliFactory$new(f_1 = list(C,1), f_2 = list(1,0))
res <- bf_logistic$sample(n = size_sample, roll.fun = toss.coin.logistic, num_cores = 2, verbose = TRUE, p=p)

mean_Huber <- C/(1+C*p)
print(mean_Huber) #Mean of Huber Algorithm
print((1+C)/(1+C*p))  #Mean of Dice Enterprise
print(mean(res[[2]]))

#CHECK IF IT IS CORRECT
table(res[[1]])/size_sample
print(c(C*p/(1+C*p), 1/(1+C*p)))

```
```{r bf_Huber_A, echo=FALSE,eval=FALSE}
#Algorithm A of Huber for function 
#f(p) = Cp(1-(Cp)^(m-1))/(1-(Cp)^m) where m=5
m <- 5
bf_A <- BernoulliFactory$new(f_1 = list(coeff = 10^seq(m,2), power = seq(m-1,1)),
                             f_2 = list(coeff = c(10), power = c(0)))

size_sample <- 10000
res_A <- bf_A$sample(n = size_sample, roll.fun = toss.coin.logistic, num_cores = 2, verbose = TRUE, p=p)

mean_Huber_A <- C*(m-1)
print(mean_Huber_A) #Mean of Huber Algorithm
print(mean(res_A[[2]])) #Empirical mean of dice enterprise

#Check correctness
table(res_A[[1]])/size_sample
print(bf_A$evaluate(p))
```

## Dice Enterprise
The procedure is similar to the previous. Assume that the given function $f(\boldsymbol{p}): \Delta^m \rightarrow \Delta^v$ is 
$$
f(p_1,p_2,p_3) \propto (p_1p_2^2+\sqrt{2}p_2^2p_3^2, 4p_1^4p_2^7+1/2+3p_1p_2^2p_3^3,7p_1p_2^3p_3^4+2p_3^2)
$$
The user needs to define a list that specify the polymomial $f(\boldsymbol{p})$. Each element corresponds to $f_i(\boldsymbol{p})$ and each of this function is determined by a list itself containing two elements:
1. A vector of the coefficients (as in the Bernoulli Factory case)
2. A matrix or a vector of string defining the power of the variables. If a matrix is given, it has to have $m$ columns and each row defines the power of $p_1,p_2,\ldots,p_m$ respectively. For instance $p_1p_2^2+\sqrt{2}p_2^2p_3^2$ is translated into the matrix 
$$
\begin{bmatrix}
1 & 2 & 0 \\
0 & 2 & 2
\end{bmatrix}
$$
Alternatively, the same powers can be defined via a vector of strings. The powers of the previous polynomial $p_1p_2^2+\sqrt{2}p_2^2p_3^2$ can also be defined as `c("120","022")`. Notice that this method works **only** if there are no powers greater than 10.

The polynomial $f(\boldsymbol{p})$ is then defined as
```{r de_polynomials}
f_dice <- list(
  list(c(1,sqrt(2)),c("120","022")),
  list(c(4,1/2,3),c("470","000","123")),
  list(c(7,2),matrix(c(1,3,4,0,0,2),byrow=TRUE,ncol=3))
)
```
To construct a Dice Enterprise with the defined polynomial the following command is run: 
```{r de_construction}
de <- DiceEnterprise$new(f_dice)
```
and an $f(\boldsymbol{p})$-die is rolled by giving a function that rolls the original $m$-sided die and using the `sample` method. 
```{r de_roll}
roll_die <- function(n) {
  sample(1:3, size = n, replace = TRUE, prob = c(1/5,1/4,1-1/5-1/4))
} #The original die has probability 1/5, 1/4, 11/20
sample_die <- de$sample(n = 10, roll.fun = roll_die)
print(sample_die)
```

### Diagnosis and extra options

Analogously as described in the Bernoulli Factory case, the method `sample` accepts the same inputs as in the Bernoulli Factory case. The function `plot.confidence.interval` returns a plot of the estimates together with a 95% confidence interval.
```{r de_diagnosis}
sample_die <- de$sample(n = 1000, roll.fun = roll_die, num_cores = 2, verbose = TRUE, double_time = TRUE) #Produces 1000 rolls of the f(p)-die, using 2 cores
print(table(sample_die[[1]])/1000) #Empirical probabilities. Notice that the theoretical ones are given by print(de$evaluate(c(1/5,1/4,1-1/5-1/4)))
print(de$evaluate(c(1/5,1/4,1-1/5-1/4)))
print(paste0("Average number of rolls required: ", mean(sample_die[[2]])))
plot.confidence.interval(sample_die[[1]],de$evaluate(c(1/5,1/4,1-1/5-1/4)))

```

### Example from the paper
This reproduces example 2 from the paper and simulates from the multivariate ladder 
$$
\pi(p_1,p_2,p_3) \propto \left(\sqrt{2}p_1^3,p_1^2p_3,\frac{1}{4}p_1p_2^2,2p_1p_2p_3,\frac{1}{2}p_1p_3^2,\frac{3}{4}p_2^2p_3\right)
$$

```{r example_2_paper}
de_ex2 <- DiceEnterprise$new(G=list(
  list(sqrt(2),"300"),
  list(1,"201"),
  list(1/4,"120"),
  list(2,"111"),
  list(1/2,"102"),
  list(3/4,"021")
))
#Define the original die
true_prob_original_die <- c(1/5,1/4,11/20)
roll_die <- function(n) {
  sample(1:3, size = n, replace = TRUE, prob = true_prob_original_die)
} 
#Get a sample of size 1000 from the multivariate ladder
#and plot the estimates with confidence intervals
sample_ex2 <- de_ex2$sample(n=1000, roll.fun = roll_die)
plot.confidence.interval(sample_ex2,de_ex2$evaluate(true_prob_original_die))

```

## Independent coins

Assume that instead of an $m$-sided die, $m$ independent coins are given, each one with its own probability of landing heads and denoted by $p_i$. In general, $(p_1,\ldots,p_m) \not\in \Delta^m$. Consider as an example, that we have access to 3 independent coins of biases $(0.4,0.7,0.55)$ and wish to sample from the distribution $f(\boldsymbol{p}) = \frac{1}{p_1+p_2+p_3}\left(p_1,p_2,p_3\right) = (8/33,14/33,1/3)$. There are several ways to define a die whose probabilites are a function of $(p_1,\ldots,p_m)$. The package provides support for two different definitions:

### Scenario 1: tossing all the coins
Considering constructing an $(m+2)$-sided die with the following probabilities of rolling $\{1,2\ldots,m+2\}$.
$$
\begin{split}
&q_1 = p_1\cdot p_2\cdot\ldots \cdot p_m \\
&q_2 = (1-p_1)\cdot p_2 \cdot \ldots \cdot p_m \\
&q_3 = p_1 \cdot (1-p_2) \cdot p_3 \cdot \ldots \cdot p_m \\
&\ldots \\
&q_{m+1} = p_1 \cdot p_2 \cdot \ldots \cdot p_{m-1} \cdot (1-p_m) \\
&q_{m+2} = 1-\sum_{i=0}^m q_i
\end{split}
$$
It is straightforward to roll such die by tossing all the $m$ independent coins at the same time. In particular, if all the tosses result in heads, then 1 is returned. If all the tosses are heads, except for the $l$th coin, then $l+1$ is returned. In all other cases, $m+2$ is returned. Finally, assume that $f(\boldsymbol{p}): (0,1)^m \rightarrow \Delta^v$ is a function of the original probabilities. It is possible to construct a function $\tilde{f}(\boldsymbol{q}): \Delta^{m+2} \rightarrow \Delta^v$ by substituing:
$$
p_i = \frac{q_1}{q_1+q_{i+1}}, \qquad i \in \{1,2,\ldots,m\}
$$

The user needs to first define a function that tosses all the three coins. Notice again that heads corresponds to 1 and tails to 2 (**not to 0**).

```{r ce_toss_all}
toss.all.coins <- function(probs) { 
  return(sapply(probs, function(p) {sample(1:2, size = 1, prob = c(p,1-p))})) #1 or 2 (not 0))
}
```

The function $f(\boldsymbol{p})$ can be expressed as a function of the transformed variables $\boldsymbol{q} = (q_1,\ldots,q_5)$ previously described as:
$$
\begin{split}
f(\boldsymbol{q}) &= \frac{1}{C(\boldsymbol{q})}\left((q_0+q_2)(q_0+q_3),(q_0+q_1)(q_0+q_3),(q_0+q_1)(q_0+q_2)\right) \\
&= \frac{1}{C(\boldsymbol{q})}\left(q_0^2+q_0q_2+q_0q_3+q_2q_3,q_0^2+q_0q_1+q_0q_3+q_1q_3,q_0^2+q_0q_1+q_0q_2+q_1q_2\right)
\end{split}
$$
where $C(\boldsymbol{q}) = 3(q_0+q_1)(q_0+q_2)(q_0+q_3)\sum_{i=1}^3 \frac{1}{q_0+q_i}$. As before, the user needs to define the three polynomials as a list of coefficients and powers:
```{r ce_toss_all_poly}
f_indep_coins1 <- list(
  list(rep(1,4),c("20000","10100","10010","00110")),
  list(rep(1,4),c("20000","11000","10010","01010")),
  list(rep(1,4),c("20000","11000","10100","01100"))
)
```

The package provides a class `CoinsEnterprise` to deal with this class of problems and automatically define the corresponding die. In particular, when an object of such class is initialized the argument `die_type` specifies which type of construction is considered. For scenario 1 this corresponds to `die_type = "toss_all"`.

```{r ce_toss_all_initialize}
ce1 <- CoinsEnterprise$new(f_indep_coins1, toss.coins = toss.all.coins, num_coins = 3, die_type = "toss_all")
```

Getting a sample from the newly create die is done analogously as before. Notice that the number of rolls does not correspond to the number of tosses of the coins. In particular, the number of tosses is obtained by multiplying the number of rolls by $m = 3$.
```{r ce_toss_all_sample}
indep_coins_probs <- c(0.4,0.7,0.55)
result <- ce1$sample(n = 1000, num_cores = 2, verbose = TRUE,  double_time = FALSE, probs = indep_coins_probs) #the argument probs is passed to toss.coins
print(table(result[[1]])/1000) #Empirical probabilities. Notice that the theoretical ones are given by
print(indep_coins_probs/sum(indep_coins_probs))
print(paste0("Average number of rolls required: ", mean(result[[2]])))
print(paste0("Average number of tosses required: ", length(indep_coins_probs)*mean(result[[2]])))
plot.confidence.interval(result[[1]],indep_coins_probs/sum(indep_coins_probs))

```

### Scenario 2: first heads
Consider constructing an $(m+1)$-sided die with the following probabilities of rolling $\{1,2,\ldots,m+1\}$.
$$
\begin{split}
q_1 &= p_1 \\
q_2 &= (1-p_1)\cdot p_2 \\
q_3 &= (1-p_1)\cdot (1-p_2)\cdot p_3 \\
&\ldots \\
q_m &= (1-p_1)\cdot (1-p_2)\cdot \ldots\cdot  (1-p_{m-1})\cdot p_m \\
q_{m+1} &= (1-p_1)\cdot (1-p_2)\cdot \ldots\cdot  (1-p_{m-1})\cdot (1-p_m)
\end{split}
$$
It is straightforward to roll such die by flipping in order the $m$ independent coins until the first heads is returned at the $l$th toss. In this case, $l$ is returned as a result of rolling the die. If all the tosses returns tails, then $l+1$ is returned. 
If $f(\boldsymbol{p}): (0,1)^m \rightarrow \Delta^v$ is a function of the original probabilities, it is possible to construct a function $\tilde{f}(\boldsymbol{q}): \Delta^{m+2} \rightarrow \Delta^v$ by substituing:
$$
p_i = \frac{q_i}{1-\sum_{k=1}^{i-1}q_k}, \qquad i \in \{1,2,\ldots,m\}
$$

The user needs to first define a function that tosses all the three coins until a heads is obtained. Notice again that heads corresponds to 1 and tails to 2 (**not to 0**).

```{r ce_toss_until_heads}
toss.until.heads <- function(probs) {
  m <- length(probs)
  res <- rep(NA, m)
  for(i in 1:m) {
    res[i] <- sample(1:2, size = 1, prob = c(probs[i],1-probs[i]))
    if(res[i] == 1) { break } #Stop loop if heads is obtained
  }
  return(res)
}
```

The function $f(\boldsymbol{p})$ can be expressed as a function of the transformed variables $\boldsymbol{q} = (q_1,\ldots,q_4)$ previously described as:
$$
\begin{split}
f(\boldsymbol{q}) &= \frac{1}{C(\boldsymbol{q})}\left(q_1(1-q_1)(1-q_1-q_2), q_2(1-q_1-q_2), q_3(1-q_1) \right) \\
&= \frac{1}{C(\boldsymbol{q})}\left(q_1 - 2q_1^2+q_1^3-q_1q_2+q_1^2q_2, q_2-q_1q_2-q_2^2,q_3-q_1q_3  \right)
\end{split}
$$
where $C(\boldsymbol{q}) = q_1(1-q_1)(1-q_1-q_2)+q_2(1-q_1-q_2)+q_3(1-q_1)$. As before, the user needs to define the three polynomials as a list of coefficients and powers:
```{r ce_toss_until_heads_poly}
f_indep_coins2 <- list(
  list(c(1,-2,1,-1,1),c("1000","2000","3000","1100","2100")),
  list(c(1,-1,-1),c("0100","1100","0200")),
  list(c(1,-1),c("0010","1010"))
)
```

In this case, an object of the class `CoinsEnterprise` must be initialized with argument `die_type = "first_heads"`.

```{r ce_toss_until_heads_initialize}
ce2 <- CoinsEnterprise$new(f_indep_coins2, toss.coins = toss.until.heads, num_coins = 3, die_type = "first_heads")
```

Getting a sample from the newly create die is done analogously as before. Notice that the number of rolls does not correspond to the number of tosses of the coins. In particular, the number of tosses is obtained by multiplying the number of rolls by the expected number of tosses of the function `toss.until.heads`. In this case, it is given by $p_1+2(1-p_1)p_2+3(1-p_1)(1-p_2) = 1.78$.
```{r ce_toss_until_heads_sample}
result <- ce2$sample(n = 1000, num_cores = 2, verbose = TRUE, double_time = FALSE, probs = indep_coins_probs) #the argument probs is passed to toss.coins
print(table(result[[1]])/1000) #Empirical probabilities. Notice that the theoretical ones are given by
print(indep_coins_probs/sum(indep_coins_probs))
print(paste0("Average number of rolls required: ", mean(result[[2]])))
print(paste0("Average number of tosses required: ", (indep_coins_probs[1]+
                                                       2*(1-indep_coins_probs[1])*indep_coins_probs[2]+
                                                       3*(1-indep_coins_probs[1])*(1-indep_coins_probs[2]))*
                                                       mean(result[[2]])))
plot.confidence.interval(result[[1]],indep_coins_probs/sum(indep_coins_probs))

```

### Scenario 3: uniformly

Consider constructing an $(m+1)$-sided die with the following probabilities of rolling $\{1,\ldots,m\}$.
\begin{align*}
q_1 &= \frac{1}{m}p_1 \\
q_2 &= \frac{1}{m}p_2 \\
&\ldots \\
q_m &= \frac{1}{m}p_m \\
q_{m+1} &= 1-\frac{1}{m}\sum_{i=1}^m p_i \\
\end{align*}
To roll such die it is just sufficient to select uniformly at random a coin $i$ to toss. If it lands heads, then $i$ is returned, otherwise $m+1$ is returned. Given a function $f(\boldsymbol{p}):(0,1)^m \rightarrow \Delta^v$ of the independent coins, it is straightforward to construct a function $f(\boldsymbol{q}): \Delta^{m+1} \rightarrow \Delta^v$ of the new constructe die by substituting:
$$
p_i = mq_i, \qquad i \in \{1,2,\ldots,m\}
$$
The user needs to define a function that takes as input which coin has to be tossed and return $1$ if the coin lands heads.
```{r ce_unif_toss}
toss.coins.single <- function(which_coin, probs) {
  return(sample(c(1,2), size = 1, prob = c(probs[which_coin], 1-probs[which_coin])))
}
```
The function $f(\boldsymbol{p})$ previously described can be expresses as a function of the transformed variables simply as
$$
f(\boldsymbol{q}) = \frac{1}{q_1+q_2+q_3}(q_1,q_2,q_3)
$$
As before, the user needs to define three polynomials as a list of coefficients and powers. Notice that there are $4$ faces of the new die, and not three.

```{r ce_unif_poly}
f_indep_coins3 <- list(
  list(1, "1000"),
  list(1, "0100"),
  list(1, "0010")
)
```

In this case, an object of the class `CoinsEnterprise` must be initialized with argument `die_type = "uniform"`.

```{r ce_unif_def}
ce3 <- CoinsEnterprise$new(f_indep_coins3, toss.coins = toss.coins.single, num_coins = 3, die_type = "uniform")
```

Getting a sample from the newly create die is analogous as before. In this case the number of rolls is equal to number of tosses of the coins.

```{r ce_unif_sample}
result <- ce3$sample(n = 1000, num_cores = 2, verbose = TRUE, double_time = FALSE, probs = indep_coins_probs) #the argument probs is passed to toss.coins.single
print(table(result[[1]])/1000) #Empirical probabilities. Notice that the theoretical ones are given by
print(indep_coins_probs/sum(indep_coins_probs))
print(paste0("Average number of tosses required: ", mean(result[[2]])))
plot.confidence.interval(result[[1]],indep_coins_probs/sum(indep_coins_probs))
```

```{r eliminiami, eval=FALSE,echo=FALSE}
roll.uniform <- function(n,probs_coins,probs_unif = rep(1,length(probs_coins))) {
  m <- length(probs_coins)
  res <- numeric(n)
  for(k in 1:n) {
  i <- sample(1:m, size = 1, prob = probs_unif)
  res[k] <- (sample(c(i,m+1), size = 1, prob = c(probs_coins[i],1-probs_coins[i])))
  }
  return(res)
}

indep_coins_probs <- seq(0.01,0.09,by=0.005)
f_unif <- vector("list", length = length(indep_coins_probs))
for(i in 1:length(indep_coins_probs)) {
  zeros <- rep(0, length(indep_coins_probs)+1)
  zeros[i] <- 1
  f_unif[[i]] <- list(c(1), paste0(zeros,collapse=""))
}

de_uniform <- DiceEnterprise$new(f_unif)

size_sample <- 10000
sample_coins_uniform <- de_uniform$sample(n = size_sample, roll.fun = roll.uniform, probs_coins = indep_coins_probs, verbose = TRUE)
table(sample_coins_uniform[[1]])/size_sample
mean(sample_coins_uniform[[2]])
length(indep_coins_probs)/sum(indep_coins_probs)
```

### 2018-04-11 inst/doc/PaperExamples.Rmd
Added vignettes

---
title: "Reproducing Paper Examples"
author: "Giulio Morina"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Main examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7, fig.height=4,
  cache=TRUE
)
require("DiceEnterprise")
require("pander")
require("ggplot2")
require("xtable")
set.seed(10,"L'Ecuyer-CMRG")
```

## Example 1 - Bernoulli Factory
$$
f(p) = \frac{\sqrt{2}p^3}{(\sqrt{2}-5)p^3+11p^2-9p+3}
$$
```{r bf_settings}
sample_size <- 1000 #Number of tosses
probs_vec <- c(0.01,0.1,0.25,0.5,0.75,0.9,0.99) #Vector of true probabilities of the p-coin
num_cores <- 4 #Number of cores used (parallel computing is not supported on Windows)
conf_level <- 0.95 #Confidence interval level
```

```{r bf_toy, echo=FALSE, results='asis'}
toss.coin <- function(n, p) {
  sample(1:2, size = n, replace = TRUE, prob = c(p,1-p))
}
bf <- BernoulliFactory$new(f_1 = list(coeff = c(sqrt(2)), power = c(3)), 
                           f_2 = list(coeff = c(-5,11,-9,3), power = c(3,2,1,0)))

bf.res <- function(toss.coin, bf) {
  bf_res <- matrix(NA, nrow = 6, ncol = length(probs_vec))
  rownames(bf_res) <- c("f(p)", "LowerCI", "Emp.f(p)", "UppCI", "Exp.TossesNoDoubling", "Exp.TossesDoubling")
  colnames(bf_res) <- as.character(probs_vec)
  for(i in 1:length(probs_vec)) {
    sample_res <- bf$sample(n = sample_size, roll.fun = toss.coin, p = probs_vec[i], num_cores = num_cores, verbose = TRUE, double_time = FALSE)
    sample_res_double_time <- bf$sample(n = sample_size, roll.fun = toss.coin, p = probs_vec[i], num_cores = num_cores, verbose = TRUE, double_time = TRUE)
    #Get lower and upper CI
    if(isTRUE(all.equal(sample_res[[1]],rep(1,sample_size)))) {
      #All equal to 1 (Heads)
      lower_ci <- 1
      upper_ci <- 1
    } else if(isTRUE(all.equal(sample_res[[1]],rep(2,sample_size)))) {
      #All equal to 2 (Tails)
      lower_ci <- 0
      upper_ci <- 0
    } else {
      ci <- MultinomCI(table(sample_res[[1]]),conf.level = conf_level)
      lower_ci <- ci[1,2]
      upper_ci <- ci[1,3]
    }
    bf_res[1,i] <- round(bf$evaluate(p = probs_vec[i])[1],2)
    bf_res[2,i] <- round(lower_ci,2)
    bf_res[3,i] <- round(1-mean(sample_res[[1]]-1),2)
    bf_res[4,i] <- round(upper_ci,2)
    bf_res[5,i] <- round(mean(sample_res[[2]]),2)
    bf_res[6,i] <- round(mean(sample_res_double_time[[2]]),2)
  }
  return(bf_res)
}
bf_res <- bf.res(toss.coin,bf)
pandoc.table(bf_res)
```

```{r bf_latex, eval = FALSE, echo = FALSE}
  xtable(bf_res, align = c("l|",rep("c",ncol(bf_res))))
```

## Example 2 - Logistic Bernoulli Factory

$$
f(p) = \frac{Cp}{1+Cp}
$$

```{r bf_logistic_setting}
constants <- c(0.1,1,10,100)
```

```{r bf_logistic, echo=FALSE, results='asis'}
res_logistic <- vector("list",length(constants))
counter <- 1
for(C in constants) {
bf_logistic <- BernoulliFactory$new(f_1 = list(coeff = c(C), power = c(1)), 
                           f_2 = list(coeff = c(1), power = c(0))) 
res_logistic[[counter]] <- bf.res(toss.coin,bf_logistic)
exp_tosses_Huber <- C/(1+C*probs_vec)
exp_tosses_CFTP <- (1+C)/(1+C*probs_vec)
res_logistic[[counter]] <- rbind(rep(C,ncol(res_logistic[[counter]])),res_logistic[[counter]],exp_tosses_CFTP,exp_tosses_Huber)
rownames(res_logistic[[counter]])[c(1,nrow(res_logistic[[counter]])-1,nrow(res_logistic[[counter]]))] <- c("C","Exp.TossesCFTP","Exp.TossesHuber")
counter <- counter + 1
pandoc.table(res_logistic[[counter-1]])
}

```

```{r bf_logistic2, echo = FALSE, eval = FALSE}
# An interpretation of the resulting CFTP algorithm leads to the following algorithm
# which is equivalent to Huber's algorithm in the average number of tosses required.
# This is just to check if it is correct.

bf.logistic.CFTP.smart <- function(n,C,p,toss.coin) {
  res <- numeric(n)
  tosses <- rep(0,n)
  i <- 1
  while(i <= n) {
    U <- runif(1)
    if(U <= 1/(1+C)) {
      res[i] <- 0
      i <- i+1
    } else {
      X <- toss.coin(1,p)
      tosses[i] <- tosses[i] + 1
      if(X == 1) {
        res[i] <- X
        i <- i+1
      } 
    }
  }
  return(list(res=res,tosses=tosses))
}

C <- 10
p <- 0.2
results <- bf.logistic.CFTP.smart(1000,C,p,toss.coin)
mean(results$res)
print(C*p/(1+C*p))
mean(results$tosses)
print(C/(1+C*p)) #Theoretical number of tosses required
```

## Example 3 - Dice Enterprise

$$
\pi(p_1,p_2,p_3) \propto \left(\sqrt{2}p_1^3,p_1^2p_3,\frac{1}{4}p_1p_2^2,2p_1p_2p_3,\frac{1}{2}p_1p_3^2,\frac{3}{4}p_2^2p_3\right)
$$
```{r de_setup}
sample_size <- 1000 #Number of rolls
probs_mat <- matrix(c(1/5,1/4,11/20,
                      1/10,4/10,5/10), ncol = 3, byrow = TRUE) #Matrix of true probabilities of the die
num_cores <- 4 #Number of cores used (parallel computing is not supported on Windows)
conf_level <- 0.95 #Confidence interval level
```

```{r de, eval=TRUE, echo = FALSE, result='asis'}

roll_die <- function(n, probs_die) {
  sample(1:3, size = n, replace = TRUE, prob = probs_die)
} 
de <-  DiceEnterprise$new(G=list(
  list(sqrt(2),"300"),
  list(1,"201"),
  list(1/4,"120"),
  list(2,"111"),
  list(1/2,"102"),
  list(3/4,"021")
))

de_res <- matrix(NA, nrow = 6, ncol = nrow(probs_mat))
rownames(de_res) <- c("f(p)", "LowerCI", "Emp.f(p)", "UppCI", "Exp.RollsNoDoubling", "Exp.RollsDoubling")
colnames(de_res) <- apply(probs_mat, 1, function(x) { paste0("(",paste0(x, collapse = ", "),")")} )

for(i in 1:nrow(probs_mat)) {
  de_sample_double_time <- de$sample(n=sample_size, roll.fun = roll_die, double_time = TRUE, num_cores = num_cores, verbose = TRUE, probs_die = probs_mat[i,])
  de_sample <- de$sample(n=sample_size, roll.fun = roll_die, double_time = FALSE, num_cores = num_cores, verbose = TRUE, probs_die = probs_mat[i,])
  if(length(unique(de_sample[[1]])) > 1) {
    #There are more outcomes
    ci <- MultinomCI(table(de_sample[[1]]),conf.level = conf_level)
    ci_aux <- matrix(NA, nrow = 2, ncol = 6) #6 possible outcomes of the die
    for(j in 1:nrow(ci)) {
      ci_aux[1,as.integer(rownames(ci)[j])] <- ci[j,2]
      ci_aux[2,as.integer(rownames(ci)[j])] <- ci[j,3]
    }
    de_res[2,i] <- paste0("(",paste0(round(ci_aux[1,],2), collapse = ", "),")") #Lower confidence interval
    de_res[4,i] <- paste0("(",paste0(round(ci_aux[2,],2), collapse = ", "),")")  #Uppter confidence interval
  }
  de_res[1,i] <- paste0("(",paste0(round(de$evaluate(p = probs_mat[i,]),2), collapse = ", "),")")
  de_res[3,i] <- paste0("(",paste0(round(table(de_sample[[1]])/sample_size,2), collapse = ", "),")")
  de_res[5,i] <- round(mean(de_sample[[2]]),2)
  de_res[6,i] <- round(mean(de_sample_double_time[[2]]),2)
}
pandoc.table(de_res,split.tables=Inf)
```

## Example 4 - Bernoulli race (independent coins)

See main vignette.

## Example 5 - 2-coin algorithm
$$
f(p) = \frac{c_1p_1}{c_1p_1+c_2p_2}, \qquad c_1, c_2 \geq 0
$$
When using the substitution $p_i = \frac{q_1}{q_1+q_{i+1}}, \qquad i \in \{1,2,\ldots,m\}$ (tossing all coins to produce an $m+2$ sided die), one gets
$$
f_{1}(\boldsymbol{q}) \propto (c_1q_1^2+c_1q_1q_3, c_2q_1^2+c_2q_1q_2)
$$
Instead, when substituing using $p_i = \frac{q_i}{1-\sum_{k=1}^{i-1}q_k}, \qquad i \in \{1,2,\ldots,m\}$ (tossing coins until first heads to produce an $m+1$ sided die), one gets
$$
f_2(\boldsymbol{q}) \propto (c_1q_1 -c_1q_1^2, c_2q_2)
$$

```{r 2coin_alg, eval=FALSE, echo=FALSE, results='asis'}
probs_sum <- c(1.9,1.8,1.5,2*2^(-seq(0,8))) #prob_sum = p_1+p_2 (and p_1 = p_2)
probs_sum <- c(1.8, 1.5, 1, 0.2,0.02,0.002,0.0002) 
sample_size <- 1000 #Number of tosses
num_cores <- 2 #Number of cores used (parallel computing is not supported on Windows)
conf_level <- 0.95 #Confidence interval level

indep_coin_probs <- matrix(NA, ncol = 2, nrow = length(probs_sum))
for(i in 1:length(probs_sum)) {
  indep_coin_probs[i,] <- c(probs_sum[i]/2,probs_sum[i]/2)
}
const <- matrix(1, nrow = nrow(indep_coin_probs),  ncol=2, byrow = TRUE) #c1 = 1, c2 = 1

toss.all.coins <- function(probs) {
  return(sapply(probs, function(p) {sample(1:2, size = 1, prob = c(p,1-p))})) #1 or 2 (not 0))
}
toss.until.heads <- function(probs) {
  m <- length(probs)
  res <- rep(NA, m)
  for(i in 1:m) {
    res[i] <- sample(1:2, size = 1, prob = c(probs[i],1-probs[i]))
    if(res[i] == 1) { break } #Stop loop if heads is obtained
  }
  return(res)
}

ce_2coins_all_tosses_res <- matrix(NA, nrow = 7, ncol = nrow(indep_coin_probs))
rownames(ce_2coins_all_tosses_res) <- c("q","f(p)", "LowerCI", "Emp.f(p)", "UppCI", "Exp.Tosses", "Exp.Tosses2coinsAlg")
colnames(ce_2coins_all_tosses_res) <- apply(indep_coin_probs, 1, paste, collapse = ", ")
ce_2coins_until_heads_res <- matrix(NA, nrow = 7, ncol = nrow(indep_coin_probs)) 
rownames(ce_2coins_until_heads_res) <- c("q","f(p)", "LowerCI", "Emp.f(p)", "UppCI", "Exp.Tosses", "Exp.Tosses2coinsAlg")
colnames(ce_2coins_until_heads_res) <- apply(indep_coin_probs, 1, paste, collapse = ", ")

for (i in 1:nrow(indep_coin_probs)) {
  ce_2coins_allcoins <- CoinsEnterprise$new(list(
  list(c(const[i,1],const[i,1]), c("2000","1010")),
  list(c(const[i,2],const[i,2]), c("2000","1100"))
), toss.coins = toss.all.coins, die_type = "toss_all")
  ce_2coins_until_heads <- CoinsEnterprise$new(list(
  list(c(const[i,1],-const[i,1]), c("100","200")),
  list(c(const[i,2]), c("010"))
), toss.coins = toss.until.heads, die_type = "first_heads")
  sample_2coins_allcoins <- ce_2coins_allcoins$sample(n = sample_size, num_cores = num_cores, verbose = TRUE, double_time = FALSE, probs = indep_coin_probs[i,])
  sample_2coins_untilheads <- ce_2coins_until_heads$sample(n = sample_size, num_cores = num_cores, verbose = TRUE, double_time = FALSE, probs = indep_coin_probs[i,])

  exp_2coin <- sum(const[i,])/sum(const[i,]*indep_coin_probs[i,])
  average_tosses_allcoins <- length(indep_coin_probs[i,])*mean(sample_2coins_allcoins[[2]]) #2*number of rolls
  average_tosses_untilheads <- (indep_coin_probs[i,1]+2*(1-indep_coin_probs[i,1]))*mean(sample_2coins_untilheads[[2]]) #Works only for two coins

  q_allcoins <- c(prod(indep_coin_probs[i,]), #Works only for three coins
                  (1-indep_coin_probs[i,1])*indep_coin_probs[i,2],
                  (1-indep_coin_probs[i,2])*indep_coin_probs[i,1])
  q_allcoins[4] <- 1-sum(q_allcoins)
  q_untileheads <- c(indep_coin_probs[i,1],
                     (1-indep_coin_probs[i,1])*indep_coin_probs[i,2],
                     (1-indep_coin_probs[i,1])*(1-indep_coin_probs[i,2]))

  #Fill case all tosses
  #ce_2coins_all_tosses_res[1,i] <- paste0("(",paste0(q_allcoins, collapse = ", "),")")
  ce_2coins_all_tosses_res[2,i] <- (const[i,1]*indep_coin_probs[i,1])/sum(const[i,]*indep_coin_probs[i,])
  ce_2coins_all_tosses_res[4,i] <- 1-mean(sample_2coins_allcoins[[1]]-1)
  ce_2coins_all_tosses_res[6,i] <- average_tosses_allcoins
  ce_2coins_all_tosses_res[7,i] <- exp_2coin

  #Fill case until heads
  #ce_2coins_until_heads_res[1,i] <- paste0("(",paste0(q_untileheads, collapse = ", "),")")
  ce_2coins_until_heads_res[2,i] <- (const[i,1]*indep_coin_probs[i,1])/sum(const[i,]*indep_coin_probs[i,])
  ce_2coins_until_heads_res[4,i] <- 1-mean(sample_2coins_untilheads[[1]]-1)
  ce_2coins_until_heads_res[6,i] <- average_tosses_untilheads
  ce_2coins_until_heads_res[7,i] <- exp_2coin

  #Confidence interval
  if(length(unique(sample_2coins_allcoins[[1]])) > 1) {
    #At least one of each result
    ci_allcoins <- MultinomCI(table(sample_2coins_allcoins[[1]]),conf.level = conf_level)
    ce_2coins_all_tosses_res[3,i] <- ci_allcoins[1,2]
    ce_2coins_all_tosses_res[5,i] <- ci_allcoins[1,3]
  }
  if(length(unique(sample_2coins_untilheads[[1]])) > 1) {
    #At least one of each result
    ci_until_heads <- MultinomCI(table(sample_2coins_untilheads[[1]]),conf.level = conf_level)
    ce_2coins_until_heads_res[3,i] <- ci_until_heads[1,2]
    ce_2coins_until_heads_res[5,i] <- ci_until_heads[1,3]
  }
}

pandoc.table(ce_2coins_all_tosses_res, round = 2)
pandoc.table(ce_2coins_until_heads_res, round = 2)

ggplot() + geom_line(aes(x = probs_sum, y = log(ce_2coins_until_heads_res[6,])), color = "red") + geom_point(aes(x = probs_sum, y = log(ce_2coins_until_heads_res[6,])), color = "red")  + 
  geom_line(aes(x = probs_sum, y = log(ce_2coins_until_heads_res[7,])), color = "blue") + geom_point(aes(x = probs_sum, y = log(ce_2coins_until_heads_res[7,])), color = "blue")

ggplot() + geom_line(aes(x = probs_sum, y = ce_2coins_until_heads_res[7,]/ce_2coins_until_heads_res[6,]), color = "red")
```

```{r bf_tex, eval = FALSE, echo = FALSE}
bf_tex <- rbind(ce_2coins_all_tosses_res,ce_2coins_until_heads_res)
bf_tex <- bf_tex[c(2:6,13,7),c(2,3,4,5:ncol(bf_tex))]
xtable(bf_tex, align = c("l|",rep("c",ncol(bf_tex))))
```

### 2018-04-11 inst/doc/DiceEnterprise.R

@@ -1,0 +1,237 @@
+## ----setup, include = FALSE----------------------------------------------
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  fig.width=7, fig.height=4,
+  cache=TRUE
+)
+require("DiceEnterprise")
+set.seed(10, "L'Ecuyer-CMRG") 
+
+## ----toss_coin-----------------------------------------------------------
+toss_coin <- function(n) {
+  sample(1:2, size = n, replace = TRUE, prob = c(3/4,1/4))
+}
+
+## ----bf_poly-------------------------------------------------------------
+f_1 <- list(coeff = c(sqrt(2)), power = c(3)) #f(p)
+f_2 <- list(coeff = c(-5,11,-9,3), power = c(3,2,1,0)) #1-f(p)
+
+## ----bf_construction-----------------------------------------------------
+bf <- BernoulliFactory$new(f_1 = f_1, f_2 = f_2) #f_1 = f(p), f_2 = 1-f(p)
+
+## ----bf_toss-------------------------------------------------------------
+fp_tosses <- bf$sample(n = 10, roll.fun = toss_coin) #Produces 10 tosses of the f(p)-coin
+print(fp_tosses)
+
+## ----bf_diagnosis--------------------------------------------------------
+bf <- BernoulliFactory$new(f_1 = f_1, f_2 = f_2, verbose = TRUE)
+print(bf)
+
+## ----bf_evaluate---------------------------------------------------------
+print(bf$evaluate(3/4))
+
+## ----bf_toss_2-----------------------------------------------------------
+fp_tosses <- bf$sample(n = 1000, roll.fun = toss_coin, num_cores = 2, verbose = TRUE, double_time= TRUE) #Produces 1000 tosses of the f(p)-coin, using 2 cores and doubling the time step at each iteration of CFTP.
+print(table(fp_tosses[[1]])/1000) #Empirical probabilities. Notice that the theoretical ones are given by print(bf$evaluate(3/4))
+print(paste0("Average number of tosses required: ",mean(fp_tosses[[2]])))
+
+## ----bf_plot-------------------------------------------------------------
+plot.confidence.interval(fp_tosses[[1]],print(bf$evaluate(3/4)))
+
+## ----bf_error, error=TRUE------------------------------------------------
[…]

### 2018-04-11 inst/doc/PaperExamples.R

@@ -1,0 +1,13 @@
+## ----setup, include = FALSE----------------------------------------------
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  fig.width=7, fig.height=4,
+  cache=TRUE
+)
+require("DiceEnterprise")
+require("pander")
+require("ggplot2")
+require("xtable")
+set.seed(10,"L'Ecuyer-CMRG")
+

### 2018-05-07 R/dice_enterprise.R
Minor bug fixes

@@ -208,7 +208,7 @@ DiceEnterprise <- R6::R6Class("DiceEnterprise",
                                        (in case, switch signs of the coefficients accordingly).")}
         #Remove zero coefficients
         if(any(R_list[[i]] == 0)) { #Need to check or M_list[[i]][numeric(0),] is empty!
-          M_list[[i]] <- M_list[[i]][-which(R_list[[i]] == 0),] #NOT GREAT, but it should work cause we rounded up the values of R
+          M_list[[i]] <- matrix(M_list[[i]][-which(R_list[[i]] == 0),], ncol = ncol(M_list[[i]]), byrow = FALSE) #NOT GREAT, but it should work cause we rounded up the values of R
           R_list[[i]] <- R_list[[i]][-which(R_list[[i]] == 0)]
         }
         M_rows <- M_rows + nrow(M_list[[i]])

### 2018-06-18 R/dice_enterprise.R
Corrected a major bug that prevented CFTP to work correctly in the case the minimum/maximum states were not the first/last ones. Added the possibility to increase the degree of a ladder. Added a check for efficiency condition. Added additional example that compares the efficiency of the algorithm.

@@ -43,28 +43,31 @@ BernoulliFactory <- R6::R6Class("BernoulliFactory",
 
 DiceEnterprise <- R6::R6Class("DiceEnterprise",
   public = list(
-    initialize = function(G, verbose = FALSE) {
+    initialize = function(G, d = NULL, verbose = FALSE) {
       #G is a list of lists (G_1,G_2,...,G_v). Each list must contain two element:
       #1) a vector of coefficients (can not be null)
       #2) a matrix of powers of p_1p_2...p_m OR a vector of strings [THIS CAN WORK ONLY FOR POWER UP TO 9!].
       #for instance p_1^2p_2,p_1p_3^3 becomes either M = [2,1,0;1,0,3] or ("210","103")
+      private$G_orig <- G
       if(verbose) {cat("Converting polynomials in internal representation...")}
       private$generate.G_poly(G) #Generate a list containing the polynomials
       if(verbose) {cat("DONE! \nGenerating initial ladder...")}
       private$generate.ladder_initial() #Generate the initial ladder
       if(verbose) {cat("DONE! \nConstructing a connected ladder...")}
-      aux <- private$ladder_initial$impose.connected()
+      aux <- private$ladder_initial$impose.connected(increase_degree = d)
       private$ladder_connected <- aux[["obj"]]
       private$A_initial_connected <- aux[["A"]]
       if(verbose) {cat("DONE! \nConstructing a fine and connected ladder...")}
       aux <- private$ladder_connected$impose.fineness()
       private$ladder_fine_connected <- aux[["obj"]]
       private$A_fine_connected <- aux[["A"]]
+      private$check.efficiency.condition() #Check if efficiency condition is satisfied
       if(verbose) {cat("DONE! \n")}
     },
     print = function() {
       cat("Original function: Delta^",private$m," -> Delta^",private$v,"\n",sep = "")
       cat("Fine and connected ladder: Delta^",private$m," -> Delta^",private$ladder_fine_connected$get.k(),"\n", sep="")
+      cat("Efficiency condition is satisfied: ",private$efficiency_condition,"\n", sep ="")
     },
     evaluate = function(p) {
       #This function evaluates the polynomials given fixed probabilites.
@@ -117,13 +120,88 @@ DiceEnterprise <- R6::R6Class("DiceEnterprise",
     },
     get.G.poly = function() {private$G_poly},
     get.ladder.initial = function() {private$ladder_initial$clone()},
-    get.ladder.fine.connected = function() {private$ladder_fine_connected$clone()}
[…]

### 2018-06-18 R/ladder.R

@@ -89,6 +89,10 @@ Ladder <- R6::R6Class("Ladder",
         private$is.connected() #Check connected
         private$compute.constant() #Compute constant a
         private$compute.transition.matrix() #Compute transition matrix for the update function
+        if(private$m == 2){ #Save minimum and maximum states
+          private$min_state <- which.min(private$M[,1])
+          private$max_state <- which.max(private$M[,1])
+        }
       } else {
         stop("The ladder is not valid.")
       }
@@ -163,6 +167,22 @@ Ladder <- R6::R6Class("Ladder",
       }
       return(currentState)
     },
+    debug.update.fun = function(reps,time_span,roll.fun = NULL, true_p = NULL) {
+      #This is useful to check if the update function works correctly (just for debug purposes)
+      #We just track the chain forward for n steps and then we see if it is close
+      #to the stationary distribution
+      if(is.null(roll.fun) && is.null(true_p)) {stop("Either declare roll.fun or the true probabilities.")}
+      if(is.null(roll.fun)) {
+        stopifnot(isTRUE(all.equal(1,sum(true_p))))
+        cat("True equilibrium: ",self$evaluate(true_p),"\n")
+        roll.fun <- function(n) {sample(1:private$m, size = n, replace = TRUE, prob = true_p)}
+      }
+      res <- numeric(reps)
+      for(i in 1:reps) {
+        res[i] <- self$update.fun(1,roll.fun(time_span),runif(time_span))
+      }
+      cat("Empirical equilibrium: ",table(res)/reps,"\n")
+    },
     sample = function(n,roll.fun = NULL, true_p = NULL, num_cores = 1, verbose = FALSE, global = FALSE, double_time = FALSE,...) {
       #Get a sample from the ladder using CFTP
       #If global = TRUE, uses a different update function that makes use of a global constant -> less efficient!
@@ -180,14 +200,12 @@ Ladder <- R6::R6Class("Ladder",
         if(global) {
           stop("global is not supported anymore as it is less efficient. Use global = FALSE")
           #CFTP(k = private$k, roll.fun = roll.fun, update.fun = self$update.fun.global,
-          #     monotonic = monotonic_CFTP, min = 1, max = private$k,verbose=verbose, double_time = double_time,...) #min, max are used only in monotonic case, otherwise they are ignored
[…]

### 2018-07-02 R/RcppExports.R
Bug fixes and new functionalities
NEW FUNCTIONALITIES:
- Added possibility to use Acceptance-Rejection to sample from a ladder;
- Added method to compute the theoretical bound on the expected number of tosses required by monotonic CFTP for a fixed p;
- Added method to impose efficiency of the ladder (either via condition on the coefficients or using the theoretical bound);

BUG CORRECTION:
- Corrected a major bug that prevented some functions to be correctly converted into ladders when the coefficients were negative at the beginnings;

NEW ADDITION/MISC:
- Added a C++ utility function that spe

@@ -13,3 +13,7 @@ CFTPCpp <- function(k, probs, connected, fine, P_cumsum, P_moves_list, monotonic
     .Call('_DiceEnterprise_CFTPCpp', PACKAGE = 'DiceEnterprise', k, probs, connected, fine, P_cumsum, P_moves_list, monotonic, min, max, verbose)
 }
 
+construct_discrete_simplex <- function(d, m) {
+    .Call('_DiceEnterprise_construct_discrete_simplex', PACKAGE = 'DiceEnterprise', d, m)
+}
+

### 2018-07-02 R/dice_enterprise.R

@@ -1,7 +1,7 @@
 BernoulliFactory <- R6::R6Class("BernoulliFactory",
   inherit = DiceEnterprise,
   public = list(
-    initialize = function(f_1,f_2,verbose = FALSE) {
+    initialize = function(f_1,f_2,verbose = FALSE, threshold = 1e2) {
       #f_1 and f_2 are lists, each containing
       #the coefficients of each term (as vector)
       #and a vector of powers of p for f(p) and 1-f(p).
@@ -26,7 +26,7 @@ BernoulliFactory <- R6::R6Class("BernoulliFactory",
       G[[2]][[1]] <- coeff
       G[[2]][[2]] <- matrix(c(power,rep(0,length(power))), ncol = 2, byrow = FALSE)
       if(verbose) {cat("DONE!\n")}
-      super$initialize(G=G, verbose = verbose)
+      super$initialize(G=G, verbose = verbose, threshold = threshold)
     },
     print = function() {
       cat("Bernoulli Factory\n")
@@ -43,7 +43,7 @@ BernoulliFactory <- R6::R6Class("BernoulliFactory",
 
 DiceEnterprise <- R6::R6Class("DiceEnterprise",
   public = list(
-    initialize = function(G, d = NULL, verbose = FALSE) {
+    initialize = function(G, d = NULL, verbose = FALSE, threshold = 1e2) {
       #G is a list of lists (G_1,G_2,...,G_v). Each list must contain two element:
       #1) a vector of coefficients (can not be null)
       #2) a matrix of powers of p_1p_2...p_m OR a vector of strings [THIS CAN WORK ONLY FOR POWER UP TO 9!].
@@ -52,7 +52,7 @@ DiceEnterprise <- R6::R6Class("DiceEnterprise",
       if(verbose) {cat("Converting polynomials in internal representation...")}
       private$generate.G_poly(G) #Generate a list containing the polynomials
       if(verbose) {cat("DONE! \nGenerating initial ladder...")}
-      private$generate.ladder_initial() #Generate the initial ladder
+      private$generate.ladder_initial(threshold = threshold) #Generate the initial ladder
       if(verbose) {cat("DONE! \nConstructing a connected ladder...")}
       aux <- private$ladder_initial$impose.connected(increase_degree = d)
       private$ladder_connected <- aux[["obj"]]
[…]

### 2018-07-02 R/ladder.R

@@ -106,6 +106,7 @@ Ladder <- R6::R6Class("Ladder",
       cat("Constant a = ",private$a,"\n",sep="")
     },
     get.connected = function() {private$connected},
+    get.fine = function() {private$fine},
     update.fun.global = function(i,B,U) {
       #Update function for the ladder using the global constant a (NOT EFFICIENT)
       stopifnot(private$connected, private$fine, length(B)==length(U))
@@ -182,6 +183,88 @@ Ladder <- R6::R6Class("Ladder",
         res[i] <- self$update.fun(1,roll.fun(time_span),runif(time_span))
       }
       cat("Empirical equilibrium: ",table(res)/reps,"\n")
+    },
+    sample.AR = function(n,roll.fun = NULL, true_p = NULL, num_cores = 1, verbose = FALSE,...) {
+      #Sample using accept-reject algorithm
+      if(is.na(private$fine) || !private$fine) {
+        stop("Sampling is possible only for fine ladders.")
+      }
+      #Define rolling function
+      if(is.null(roll.fun) && is.null(true_p)) {stop("Either declare roll.fun or the true probabilities.")}
+      if(is.null(roll.fun)) {
+        stopifnot(isTRUE(all.equal(1,sum(true_p))))
+        roll.fun <- function(n) {sample(1:private$m, size = n, replace = TRUE, prob = true_p)}
+      }
+
+      #Compute constant Q = max R_i/choose(d,n) where n is any member of the discrete simplex
+      discrete_simplex <- discrete.simplex(d=private$degree,m=private$m) #all possible combinations
+      discrete_simplex_binomial_unique <- unique(as.numeric(apply(discrete_simplex,1,function(vec) {multinomial.coeff(d=private$degree,n=vec)})))
+      Q <- max(sapply(private$R, function(r) {max(r/discrete_simplex_binomial_unique)}))
+      if(verbose && !is.null(true_p)) { #Compute C_p is true_p is available
+        aux <- rep(NA, nrow(private$M))
+        for(i in 1:nrow(private$M)) {
+          aux[i] <- prod(true_p^private$M[i,])*private$R[i]
+        }
+        C_p <- sum(aux)
+      } else {
+        C_p <- NA
+      }
+
+      #Sample using A-R
+      res_AR <- mclapply(1:n, function(rep) {
+        num_rolls <- 0
+        num_iter <- 0
+        while(TRUE) {
+          num_iter <- num_iter + 1
+          #Roll the die d times
+          toss_dice <- roll.fun(n = private$degree)
[…]

### 2018-07-02 R/utils.R

@@ -61,11 +61,17 @@ discrete.simplex <- function(d,m) {
   #Generates all possible vectors of length m such that the
   #sum of its entries is equal to d.
   #Returns a matrix with m columns
-  partitions <- restrictedparts(n=d,m=m,include.zero = TRUE) #all possible partitions. Still need to permutate them.
-  l_permutations <- lapply(split(t(partitions), 1:ncol(partitions)), function(vec) {
-    data.frame(unique(t(apply(permutations(n = m, r = m), 1, function(x) vec[x]))))
-  }) #list of all the permutations (by row)
-  return(as.matrix(rbindlist(l_permutations)))
+
+  #OLD:
+  # partitions <- restrictedparts(n=d,m=m,include.zero = TRUE) #all possible partitions. Still need to permutate them.
+  # l_permutations <- lapply(split(t(partitions), 1:ncol(partitions)), function(vec) {
+  #   data.frame(unique(t(apply(permutations(n = m, r = m), 1, function(x) vec[x]))))
+  # }) #list of all the permutations (by row)
+  # return(as.matrix(rbindlist(l_permutations)))
+
+  #NEW (C++ implementation, 300x faster):
+  #This function is kept for backward compatibility
+  return(construct_discrete_simplex(d,m))
 }
 
 multinomial.coeff <- function(d,n) {

### 2018-07-02 src/RcppExports.cpp

@@ -55,11 +55,24 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
+// construct_discrete_simplex
+arma::mat construct_discrete_simplex(int d, int m);
+RcppExport SEXP _DiceEnterprise_construct_discrete_simplex(SEXP dSEXP, SEXP mSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< int >::type d(dSEXP);
+    Rcpp::traits::input_parameter< int >::type m(mSEXP);
+    rcpp_result_gen = Rcpp::wrap(construct_discrete_simplex(d, m));
+    return rcpp_result_gen;
+END_RCPP
+}
 
 static const R_CallMethodDef CallEntries[] = {
     {"_DiceEnterprise_findIntervalSingle", (DL_FUNC) &_DiceEnterprise_findIntervalSingle, 2},
     {"_DiceEnterprise_updateFunCpp", (DL_FUNC) &_DiceEnterprise_updateFunCpp, 7},
     {"_DiceEnterprise_CFTPCpp", (DL_FUNC) &_DiceEnterprise_CFTPCpp, 10},
+    {"_DiceEnterprise_construct_discrete_simplex", (DL_FUNC) &_DiceEnterprise_construct_discrete_simplex, 2},
     {NULL, NULL, 0}
 };
 

### 2018-07-02 src/dice_enterprise.cpp

@@ -1,0 +1,62 @@
+// [[Rcpp::depends(RcppArmadillo)]]
+#include <RcppArmadillo.h>
+using namespace Rcpp;
+using namespace arma;
+
+// [[Rcpp::export]]
+arma::mat construct_discrete_simplex(int d, int m) {
+  //Construct a matrix with m columns where the members are
+  //integers (possibly null) and the sum of the rows is d.
+  arma::mat discrete_simplex = zeros<mat>(R::choose(d+m-1,m-1),m);
+  int i;
+  int counter = 0;
+  if(m == 1) { //There is only one element
+    discrete_simplex[0,0] = d;
+    return(discrete_simplex);
+  }
+  for(i=0; i<=d; i++) {
+    //We set the first element equal to i and all the other elements
+    //we get them recursively
+    arma::mat aux = construct_discrete_simplex(d-i,m-1);
+    discrete_simplex.submat(counter,0,counter+aux.n_rows-1,0) = i*ones<mat>(aux.n_rows,1);
+    discrete_simplex.submat(counter,1,counter+aux.n_rows-1,m-1) = aux;
+    counter = counter+aux.n_rows;
+  }
+  return(discrete_simplex);
+}
+
+// // NOT FINISHED
+// List generate_ladder_initial_Cpp(List G_poly, int degree, int m, int v) {
+//   //This function makes sure that:
+//   //1) The polynomials are homogeneous (same degree)
+//   //2) The coefficients of the polynomials are all positive
+//   //3) All the polynomials have the same degree
+//
+//   //G_poly is a list of v elements. Each element is itself a list of 3 elements
+//   //where the first elements are the coefficients R, the second element
+//   //is a matrix of powers of the p_i, the 3rd element is the degree associated
+//   //to each coefficients (it's just the sum of the rows of the second element)
+//
+//   int i, j, k;
+//
+//   for(i = 0; i<G_poly.length(); i++) {
+//     colvec R = G_poly[i][0]; //Coefficients of the polynomial
+//     mat M = G_poly[i][1]; //Powers of the polynomial
+//     ivec degree_R = G_poly[i][2]; //Degree of each coefficients
+//     for(j=0; j<R.length(); j++) {
+//       //1 STEP: Convert to homogeneous polynomial
+//       if(degree_R[j] < degree) { //The degree of the coefficient is smaller than the degree of the polynomial
+//         //We multiply the coefficient by (p_1+...+p_m)^(degree-degree_R) to get
[…]

### 2019-01-10 R/CFTP.R
Sped up the CFTP algorithm via a new implementation of the update function in C++; New and updated methods to compute bound on the expected number of tosses; New improved construction of the transition matrix that achieves Peskun optimality.

@@ -21,7 +21,10 @@
 #' @param verbose if \code{TRUE}, the time steps required by the algorithm is returned as well.
 #' @param double_time if \code{TRUE} at each iteration of the algorithm, the time is doubled instead
 #' of increased by one. Can lead to slower implementation if rolling the die is computationally costly.
-CFTP <- function(k, roll.fun, update.fun, monotonic = FALSE, min = NA, max = NA, verbose = FALSE, double_time = FALSE,...) {
+#' @param update.fun.vec update function of the Markov chain. See the method \code{update.fun.vec} of
+#' class \code{Ladder} to check how it should be defined.
+CFTP <- function(k, roll.fun, update.fun, monotonic = FALSE, min = NA, max = NA, verbose = FALSE, double_time = FALSE,
+                 update.fun.vec=NULL,...) {
   #if double_time = true, it doubles the time span at each step. Otherwise it is just increased by one.
   #if verbose = true, a list is returned containing the time required in terms of iterations
   #min, max are used only in montonoic case, otherwise they are ignored
@@ -34,32 +37,76 @@ CFTP <- function(k, roll.fun, update.fun, monotonic = FALSE, min = NA, max = NA,
   X <- X_span
   B <- roll.fun(n=1,...)
   U <- runif(n=1)
+  mapped_states <- rep(0,k)
   #Move already forward (as coalesce may occur in 1 time step)
-  counter <- 1
-  for(i in X_span) {
-    X[counter] <- update.fun(i,B,U)
-    counter <- counter + 1
+  X <- update.fun.vec(states=X_span,B=B,U=U,
+                          current_time=1,mapped_states=mapped_states,
+                          t_mapped_states=0)
+  if(monotonic) {
+    mapped_states[min] <- X[1]; mapped_states[max] <- X[2];
+  } else {
+    mapped_states <- X
   }
+
+  # counter <- 1
+  # for(i in X_span) {
+  #   X[counter] <- update.fun(i,B,U)
+  #   counter <- counter + 1
+  # }
+
   #Loop
-  while(length(unique(X)) > 1) {
-    if(double_time) {
-      B <- c(roll.fun(n=T/2,...),B) #NOT EFFICIENT
-      U <- c(runif(n=T/2),U)
+  #while(length(unique(X, MARGIN = 2)) > 1) { #SLOWER
+  while(!all(X==X[1])) {
+
+    if(monotonic) {
[…]

### 2019-01-10 R/RcppExports.R

@@ -9,11 +9,19 @@ updateFunCpp <- function(currentState, B, U, connected, fine, P_cumsum, P_moves_
     .Call('_DiceEnterprise_updateFunCpp', PACKAGE = 'DiceEnterprise', currentState, B, U, connected, fine, P_cumsum, P_moves_list)
 }
 
+updateFunVecCpp <- function(states, B, U, connected, fine, P_cumsum, P_moves_list, mapped_states, k, t_mapped_states, current_time) {
+    .Call('_DiceEnterprise_updateFunVecCpp', PACKAGE = 'DiceEnterprise', states, B, U, connected, fine, P_cumsum, P_moves_list, mapped_states, k, t_mapped_states, current_time)
+}
+
 CFTPCpp <- function(k, probs, connected, fine, P_cumsum, P_moves_list, monotonic, min, max, verbose) {
     .Call('_DiceEnterprise_CFTPCpp', PACKAGE = 'DiceEnterprise', k, probs, connected, fine, P_cumsum, P_moves_list, monotonic, min, max, verbose)
 }
 
 construct_discrete_simplex <- function(d, m) {
     .Call('_DiceEnterprise_construct_discrete_simplex', PACKAGE = 'DiceEnterprise', d, m)
+}
+
+expected_tosses_bound_cpp <- function(A) {
+    .Call('_DiceEnterprise_expected_tosses_bound_cpp', PACKAGE = 'DiceEnterprise', A)
 }
 

### 2019-01-10 R/dice_enterprise.R

@@ -62,6 +62,7 @@ DiceEnterprise <- R6::R6Class("DiceEnterprise",
       private$ladder_fine_connected <- aux[["obj"]]
       private$A_fine_connected <- aux[["A"]]
       private$check.efficiency.condition() #Check if efficiency condition is satisfied
+      private$d <- private$ladder_fine_connected$get.degree()
       if(verbose) {cat("DONE! \n")}
     },
     print = function() {
@@ -200,16 +201,27 @@ DiceEnterprise <- R6::R6Class("DiceEnterprise",
     get.ladder.initial = function() {private$ladder_initial$clone()},
     get.ladder.fine.connected = function() {private$ladder_fine_connected$clone()},
     get.efficiency.condition = function() {private$efficiency_condition},
+    expected.tosses.bound.generic = function() {
+      #This bound is independent on p
+      warning("This bound is not theoretically valid.")
+      if(private$m > 2) {
+        stop("The expected number of tosses is computable only when the original die has 2 faces.")
+      }
+      P <- private$ladder_fine_connected$get.P()
+      a <- min(P[which(P > 0)], na.rm = TRUE)
+      k <-  private$ladder_fine_connected$get.k()
+      return(private$m^(private$d-1)/(a^(private$d)))
+    },
     expected.tosses.bound = function(true_p) {
       #This method computed the expected tosses required by CFTP for a given p.
-      #Works only when m = 2.
+      #Works only when m = 2 and the efficiency condition is satisfied.
       #true_p = (p,1-p)
       if(private$m > 2) {
         stop("The expected number of tosses is computable only when the original die has 2 faces.")
       }
-      # if(!private$efficiency_condition) {
[…]

### 2019-01-10 R/ladder.R

@@ -88,7 +88,9 @@ Ladder <- R6::R6Class("Ladder",
         private$define.neighbourhoods() #Compute neighbourhoods
         private$is.connected() #Check connected
         private$compute.constant() #Compute constant a
-        private$compute.transition.matrix() #Compute transition matrix for the update function
+        if(private$connected && private$fine) {
+          private$compute.transition.matrix() #Compute transition matrix for the update function
+        }
         if(private$m == 2){ #Save minimum and maximum states
           private$min_state <- which.min(private$M[,1])
           private$max_state <- which.max(private$M[,1])
@@ -145,6 +147,11 @@ Ladder <- R6::R6Class("Ladder",
     update.fun = function(i,B,U) {
       return(updateFunCpp(currentState = i,B = B,U = U, connected = private$connected, fine = private$fine,
                           P_cumsum = private$P_cumsum, P_moves_list = private$P_moves_list))
+    },
+    update.fun.vec = function(states,B,U,current_time,mapped_states=rep(0,private$k),t_mapped_states=0) {
+      return(updateFunVecCpp(states = states,B = B,U = U, connected = private$connected, fine = private$fine,
+                          P_cumsum = private$P_cumsum, P_moves_list = private$P_moves_list,
+                          mapped_states = mapped_states, k=private$k, t_mapped_states=t_mapped_states, current_time = current_time))
     },
     update.fun.slow = function(i,B,U) {
       #Update function for the ladder using local moves (more efficient than the global bersion)
@@ -286,7 +293,8 @@ Ladder <- R6::R6Class("Ladder",
           #     monotonic = monotonic_CFTP, min = private$min_state, max = private$max_state,verbose=verbose, double_time = double_time,...) #min, max are used only in monotonic case, otherwise they are ignored
         } else {
           CFTP(k = private$k, roll.fun = roll.fun, update.fun = self$update.fun,
-               monotonic = monotonic_CFTP, min = private$min_state, max = private$max_state,verbose=verbose, double_time = double_time,...) #min, max are used only in monotonic case, otherwise they are ignored
+               monotonic = monotonic_CFTP, min = private$min_state, max = private$max_state,verbose=verbose, double_time = double_time,
[…]

### 2019-01-10 src/CFTP_updateFun.cpp

@@ -2,6 +2,7 @@
 //#include <Rcpp.h>
 #include <RcppArmadilloExtensions/sample.h>
 using namespace Rcpp;
+using namespace std;
 
 // [[Rcpp::export]]
 int findIntervalSingle(double U, NumericVector breaks) {
@@ -14,9 +15,11 @@ int findIntervalSingle(double U, NumericVector breaks) {
   return(out);
 }
 
+
 // [[Rcpp::export]]
-int updateFunCpp(int currentState, arma::uvec B, arma::vec U, bool connected, bool fine,
-                           List P_cumsum, List P_moves_list) {
+int updateFunCpp(int currentState, const arma::uvec& B,
+                 const arma::vec& U, bool connected, bool fine,
+                 const List& P_cumsum, const List& P_moves_list) {
   //Check if it is a valid ladder and a valid request
   if(!connected || !fine || B.n_elem!= U.n_elem) {
     stop("Not a valid call. Check if the ladder is fine and connected and that B and U have the same length.");
@@ -44,6 +47,47 @@ int updateFunCpp(int currentState, arma::uvec B, arma::vec U, bool connected, bo
   }
 
   return(currentState);
+}
+
+// [[Rcpp::export]]
+arma::rowvec updateFunVecCpp(const arma::vec& states, const arma::uvec& B,
+                             const arma::vec& U, bool connected, bool fine,
+                          const List& P_cumsum, const List& P_moves_list,
+                          const arma::vec& mapped_states, int k, int t_mapped_states,
+                          int current_time) {
+  // We reuse some previous results. mapped_states is a k-long vector where the first entry
+  // tells where the first state is mapped at time 0 starting from time -t_mapped_states;
+  // the second entry tells where the second state is mapped at time 0 starting from time -t_mapped_states
+  // and so on. Some entries may not be available, in this case we need to compute it.
+
+  //Check if it is a valid ladder and a valid request
+  if(!connected || !fine || B.n_elem!= U.n_elem) {
+    stop("Not a valid call. Check if the ladder is fine and connected and that B and U have the same length.");
+  }
+
+  arma::rowvec current_states(states.n_elem);
+  int new_state;
+
+  for(int j=0; j<states.n_elem; j++) {
+    // Move state[j] from current_time to t_mapped_states
+    arma::vec U1 = U.head(current_time-t_mapped_states);
[…]

### 2019-01-10 src/RcppExports.cpp

@@ -19,19 +19,40 @@ BEGIN_RCPP
 END_RCPP
 }
 // updateFunCpp
-int updateFunCpp(int currentState, arma::uvec B, arma::vec U, bool connected, bool fine, List P_cumsum, List P_moves_list);
+int updateFunCpp(int currentState, const arma::uvec& B, const arma::vec& U, bool connected, bool fine, const List& P_cumsum, const List& P_moves_list);
 RcppExport SEXP _DiceEnterprise_updateFunCpp(SEXP currentStateSEXP, SEXP BSEXP, SEXP USEXP, SEXP connectedSEXP, SEXP fineSEXP, SEXP P_cumsumSEXP, SEXP P_moves_listSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< int >::type currentState(currentStateSEXP);
-    Rcpp::traits::input_parameter< arma::uvec >::type B(BSEXP);
-    Rcpp::traits::input_parameter< arma::vec >::type U(USEXP);
+    Rcpp::traits::input_parameter< const arma::uvec& >::type B(BSEXP);
+    Rcpp::traits::input_parameter< const arma::vec& >::type U(USEXP);
     Rcpp::traits::input_parameter< bool >::type connected(connectedSEXP);
     Rcpp::traits::input_parameter< bool >::type fine(fineSEXP);
-    Rcpp::traits::input_parameter< List >::type P_cumsum(P_cumsumSEXP);
-    Rcpp::traits::input_parameter< List >::type P_moves_list(P_moves_listSEXP);
+    Rcpp::traits::input_parameter< const List& >::type P_cumsum(P_cumsumSEXP);
+    Rcpp::traits::input_parameter< const List& >::type P_moves_list(P_moves_listSEXP);
     rcpp_result_gen = Rcpp::wrap(updateFunCpp(currentState, B, U, connected, fine, P_cumsum, P_moves_list));
+    return rcpp_result_gen;
+END_RCPP
+}
+// updateFunVecCpp
[…]

### 2019-01-10 src/dice_enterprise.cpp

@@ -25,6 +25,28 @@ arma::mat construct_discrete_simplex(int d, int m) {
   return(discrete_simplex);
 }
 
+// [[Rcpp::export]]
+double expected_tosses_bound_cpp(arma::mat A) {
+  //Perform eigen decomposition
+  arma::cx_vec eigval;
+  arma::cx_mat eigvec;
+  arma::eig_gen(eigval, eigvec, A);
+
+  arma::mat Q = real(eigvec);
+  arma::colvec D_vec = real(eigval);
+  arma::mat Q_inv = inv(Q);
+
+  //Compute expected number of tosses
+  double expected_tosses = 0.0, new_expected_tosses = 1.0;
+  int t = 1;
+  while(new_expected_tosses - expected_tosses > 1e-4) {
+    if(t > 1) {expected_tosses = new_expected_tosses;}
+    new_expected_tosses = expected_tosses + accu(Q*diagmat(pow(D_vec,t))*Q_inv);
+    t++;
+  }
+  return(new_expected_tosses);
+}
+
 // // NOT FINISHED
 // List generate_ladder_initial_Cpp(List G_poly, int degree, int m, int v) {
 //   //This function makes sure that:

### 2019-01-15 inst/doc/DiceEnterprise.Rmd
Update vignettes

@@
 $$
-f(\boldsymbol{p}) = f(p_1,\ldots,p_m) = \frac{1}{C(\boldsymbol{p})}\left(G_1(\boldsymbol{p},\ldots,G_v(\boldsymbol{p})\right)
+f(\boldsymbol{p}) = f(p_1,\ldots,p_m) = \frac{1}{C(\boldsymbol{p})}\left(G_1(\boldsymbol{p}),\ldots,G_v(\boldsymbol{p})\right)
 $$
@@
 
-The user can define a function that tosses a $p$-coin and returns $1$ with probability $p$ and 2 with probability $1-p$ (notice that **it should not return 1 or 0**).
+The user can define a function that tosses a $p$-coin and returns $1$ with probability $p$ and 2 with probability $1-p$ (notice that it **must** return 1 and 2).
 The function should have as input the number `n` of required tosses.
@@
 
-### Example: Logistic Bernoulli Factory
+When the function reaches close 1 or 0 within the interval $p \in (0,1)$, the construction of the ladder is computationally difficult and a runtime error may occur.
+Take for instance $f(p) = (x-1/2)^2 + 0.999$.
+The user can increase the timeout threshold by specifying the parameter threshold when constructing the BernoulliFactory object (default 100 iterations).
 
-Consider the function as in Huber
-$$
-f(p) = \frac{Cp}{1+Cp}
-$$
-Huber proposes a very clever algorithm that makes use of a thinned Poisson point process and requires on average $C/(1+Cp)$ tosses.
-If my calculations are correct, the Dice Enterprise can construct a Bernoulli Factory that albeit being always worse, requires at most one extra toss on average.
-Let's check this:
+```{r bf_threshold, error=TRUE}
+bf_error <- BernoulliFactory$new(f_1 = list(c(-1,1,0.749),c(2,1,0)), 
+                                 f_2 = list(c(1,-1,0.251),c(2,1,0)))
+bf_no_error <- BernoulliFactory$new(f_1 = list(c(-1,1,0.749),c(2,1,0)), 
+                                 f_2 = list(c(1,-1,0.251),c(2,1,0)),
+                                 threshold = 500)
+```
 
-```{r bf_logistic}
-toss.coin.logistic <- function(n,p) {
-  sample(1:2, size = n, replace = TRUE, prob = c(p,1-p))
-}
-C <- 10
-p <- 0.01
-size_sample <- 10000
+<!--
+### Alternative sampling algorithm
 
-bf_logistic <- BernoulliFactory$new(f_1 = list(C,1), f_2 = list(1,0))
-res <- bf_logistic$sample(n = size_sample, roll.fun = toss.coin.logistic, num_cores = 2, verbose = TRUE, p=p)
+For Bernoulli Factory type problems (where $m = 2$) it is possible to use an alternative algorithm to get a sample based on acceptance-rejection.
 
-mean_Huber <- C/(1+C*p)
-print(mean_Huber) #Mean of Huber Algorithm
-print((1+C)/(1+C*p))  #Mean of Dice Enterprise
-print(mean(res[[2]]))
+```{r bf_AR}
+fp_tosses_AR <- bf$sample.AR(n = 1000, roll.fun = toss_coin)
+plot.confidence.interval(fp_tosses_AR,bf$evaluate(3/4))
+```
+By setting the option `verbose = TRUE`, the produced output is a list that contains also further information. In particular the empirical number of tosses required by the algorithm can be computed.
+```{r bf_AR_verbose1}
+fp_tosses_AR2 <- bf$sample.AR(n = 1000, roll.fun = toss_coin, verbose = TRUE)
+print(paste0("Empirical number of tosses required: ",mean(fp_tosses_AR2$empirical_tosses)))
+```
 
-#CHECK IF IT IS CORRECT
-table(res[[1]])/size_sample
-print(c(C*p/(1+C*p), 1/(1+C*p)))
+If instead of a function that tosses the original coin, the true value of $p$ is passed to the function and `verbose = TRUE`, further information are available.
+In particular, it is possible to compute the exact expected number of tosses required.
 
+```{r bf_AR_verbose2}
+fp_tosses_AR3 <- bf$sample.AR(n = 1000, true_p = c(3/4,1/4), verbose = TRUE)
+print(paste0("Theoretical number of tosses required: ",mean(fp_tosses_AR3$theor_tosses)))
 ```
-```{r bf_Huber_A, echo=FALSE,eval=FALSE}
-#Algorithm A of Huber for function 
-#f(p) = Cp(1-(Cp)^(m-1))/(1-(Cp)^m) where m=5
-m <- 5
-bf_A <- BernoulliFactory$new(f_1 = list(coeff = 10^seq(m,2), power = seq(m-1,1)),
-                             f_2 = list(coeff = c(10), power = c(0)))
-
-size_sample <- 10000
-res_A <- bf_A$sample(n = size_sample, roll.fun = toss.coin.logistic, num_cores = 2, verbose = TRUE, p=p)
-
-mean_Huber_A <- C*(m-1)
-print(mean_Huber_A) #Mean of Huber Algorithm
-print(mean(res_A[[2]])) #Empirical mean of dice enterprise
-
-#Check correctness
-table(res_A[[1]])/size_sample
-print(bf_A$evaluate(p))
-```
+-->
 
@@
 Each element corresponds to $f_i(\boldsymbol{p})$ and each of this function is determined by a list itself containing two elements:
+
 1. A vector of the coefficients (as in the Bernoulli Factory case)
-2.
-A matrix or a vector of string defining the power of the variables.
-If a matrix is given, it has to have $m$ columns and each row defines the power of $p_1,p_2,\ldots,p_m$ respectively.
+2. A matrix or a vector of string defining the power of the variables. If a matrix is given, it has to have $m$ columns and each row defines the power of $p_1,p_2,\ldots,p_m$ respectively.
 For instance $p_1p_2^2+\sqrt{2}p_2^2p_3^2$ is translated into the matrix 
@@
 #and plot the estimates with confidence intervals
-sample_ex2 <- de_ex2$sample(n=1000, roll.fun = roll_die)
-plot.confidence.interval(sample_ex2,de_ex2$evaluate(true_prob_original_die))
+set.seed(17)
+sample_ex2 <- de_ex2$sample(n=1000, roll.fun = roll_die, verbose = TRUE)
+print(paste0("Average number of rolls required: ", mean(sample_ex2[[2]])))
+plot.confidence.interval(sample_ex2[[1]],de_ex2$evaluate(true_prob_original_die))
 
@@
 There are several ways to define a die whose probabilites are a function of $(p_1,\ldots,p_m)$.
-The package provides support for two different definitions:
+The package provides native support for three different definitions:
 
@@
 ```
+
+```{r bug_to_correct, error=TRUE, echo = FALSE}
+#(0.2501 - 1 x + 1 x^2)/((0.3 - 1 x + 1 x^2)+(0.2501 - 1 x + 1 x^2))
+# bf_bug <- BernoulliFactory$new(
+#   f_1 = list(c(0.2501,-1,1), c(0,1,2)),
+#   f_2 = list(c(0.3,-1,1), c(0,1,2)),
+# )
+#(2-5x+5x^2)/(10x+10)
+bf_bug <- DiceEnterprise$new(G = list(
+  list(c(2,-5,5),c("00","10","20")),
+  list(c(10,10),c("10","00"))
+))
+```

### 2019-01-15 inst/doc/Efficiency.Rmd
Update vignettes

---
title: "Efficiency"
author: "Giulio Morina"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7, fig.height=4,
  cache=TRUE
)
require("DiceEnterprise")
library(plyr)
```

## Example 1
We consider a ladder of the following form:
$$
\pi(p) = \frac{1}{p^a+(1-p)^a} ((1-p)^a, p^a)
$$
for different values of $a$. Notice that the denominator has a minimum in $p = 1/2$ resulting in $C(1/2) = 2^{1-a}$. We then expect Acceptance-Rejection to perform poorly as $a$ increases and $p \approx 1/2$.

We compare three methods:

1. Acceptance-Rejection: we use the theoretical value of the expecte number of tosses;

2. Naive CFTP: the proposed method targeting directly the $\pi(p)$ defined above. We run it 1,000 times;

3. Optimised CFTP: we first get an estimate of the true $p$ by tossing the $p$-coin 1,000 times and computing the sample average. We then increase the degree of the ladder using the theoretical bound on the expected number of tosses computed using such estimate of $p$. We then run CFTP 1,000 times on this augmented ladder.

```{r setup_efficiency, echo = FALSE}
p_seq <- c(0.01,0.1,0.25,0.5) #Sequence of true values of p
a_seq <- 1:10 #Sequence of a
size_sample_CFTP <- 1000 #Size sample CFTP
size_sample_est_p <- 1000 #Size tosses to get an estimate of p
num_cores <- 2 #Number of cores used
initial_seed <- 17 #Seed for RNG

generate.data <- function() {
  set.seed(initial_seed, "L'Ecuyer-CMRG") 
  data_out <- vector("list", length = length(a_seq))
  counter <- 1
  for(a in a_seq) {
    cat("Started a = ",a,"\n")
    if(file.exists(paste0("efficiency_a_",a,".rds"))) {
      #Recover previous work
      cat("Recover previous work for a = ",a,"\n")
      work_prev <- readRDS(paste0("efficiency_a_",a,".rds"))
      data_out[[counter]] <- work_prev$data_out
      counter <- counter+ 1
      seed <- work_prev$seed
      set.seed(seed, "L'Ecuyer-CMRG")  #Restore random seed
    } else {
      #Construct the ladder
      cat("Constructing the ladder\n")
      de <- DiceEnterprise$new(G=list(
        list(1,matrix(c(a,0),nrow=1,ncol=2)),
        list(1,matrix(c(0,a),nrow=1,ncol=2))
      ))
      #Get expected value for A-R
      cat("Get expected tosses of AR\n")
      exp_tosses_AR <- sapply(p_seq, function(p) {
        C_p = p^a+(1-p)^a
        Q <- 1
        return(Q*a/C_p)
      })
      cat("Get empirical expected tosses of naive CFTP\n")
      #Get value for naive CFTP
      tosses_CFTP <- lapply(p_seq, function(p) {
        return(1)
        return(de$sample(n=size_sample_CFTP, true_p = c(p,1-p), num_cores = num_cores, verbose = TRUE)[[2]])
      })
      cat("Get empirical expected tosses of augmented CFTP\n")
      #Get an estimate of p
      p_est_seq <- sapply(p_seq, function(p) {
        mean(sample(c(1,0), prob = c(p,1-p), size = size_sample_est_p, replace = TRUE))
      })
      #Get a value for augmented CFTP
      tosses_augmented_CFTP <- lapply(1:length(p_est_seq), function(i){
        p_est <- p_est_seq[i]
        p <- p_seq[i]
        #Construct augmented ladder
        new_de <- de$impose.efficiency(method="bound",true_p = c(p_est,1-p_est))
        saveRDS(list(new_de=new_de,a=a,p_est=p_est,p=p), file = paste0("augmented_de_a_",a,"_pest_",p_est,".rds"))
        return(new_de$sample(n=size_sample_CFTP, true_p = c(p,1-p), num_cores = num_cores, verbose = TRUE)[[2]])
      })

      #Prepare output
      data_out[[counter]] <- list(a = a,
                                p_seq = p_seq,
                                tosses_CFTP = tosses_CFTP,
                                exp_tosses_AR = exp_tosses_AR,
                                tosses_augmented_CFTP = tosses_augmented_CFTP) 
      #Generate new random seed
      seed <- ceiling(runif(1)*1e8)
      set.seed(seed, "L'Ecuyer-CMRG")

      #Save to outer file
      cat("Saving result for a = ",a,"\n")
      saveRDS(list(data_out=data_out[[counter]],seed=seed), file = paste0("efficiency_a_",a,".rds"))
      counter <- counter + 1
    }
  }
  return(data_out)
}
generate.data.frame <- function(data_out) {
  #Prepare output
  tot_rows <- length(a_seq)*length(p_seq)*3
  res <- data.frame(p=rep(NA,tot_rows),
                    a=rep(NA,tot_rows), 
                    meanTosses=rep(NA,tot_rows),
                    maxMeanTosses=rep(NA,tot_rows),
                    minMeanTosses=rep(NA,tot_rows),
                    method=rep(NA,tot_rows))
  counter <- 1
  for(i in 1:length(data_out)) {
    a <- data_out[[i]]$a
    p_seq_saved <- data_out[[i]]$p_seq
    exp_tosses_AR <- data_out[[i]]$exp_tosses_AR
    exp_tosses_CFTP <- sapply( data_out[[i]]$tosses_CFTP, mean)
    max_exp_tosses_CFTP <- sapply( data_out[[i]]$tosses_CFTP, max)
    min_exp_tosses_CFTP <- sapply( data_out[[i]]$tosses_CFTP, min)
    exp_tosses_augmented_CFTP <- sapply( data_out[[i]]$tosses_augmented_CFTP, mean)
    max_exp_tosses_augmented_CFTP <- sapply( data_out[[i]]$tosses_augmented_CFTP, max)
    min_exp_tosses_augmented_CFTP <- sapply( data_out[[i]]$tosses_augmented_CFTP, min)

    #Sanity check
    if(!(a %in% a_seq)) {
      stop("There's a conflict between the setup and saved results (a is not in a_seq).")
    }
    if(!isTRUE(all.equal(p_seq_saved,p_seq))) {
      stop("There's a conflict between the setup and saved results (saved p_seq is not equal to p_seq).")
    }
    #Save to data frame
    for(k in 1:length(exp_tosses_CFTP)) {
      res[counter,] <- c(p_seq_saved[k],a,exp_tosses_CFTP[k],max_exp_tosses_CFTP[k],min_exp_tosses_CFTP[k],1)
      res[counter+1,] <- c(p_seq_saved[k],a,exp_tosses_augmented_CFTP[k],max_exp_tosses_augmented_CFTP[k],min_exp_tosses_augmented_CFTP[k],2)
      res[counter+2,] <- c(p_seq_saved[k],a,exp_tosses_AR[k],exp_tosses_AR[k],exp_tosses_AR[k],3)
      counter <- counter + 3
    }
  }
  res$method <- as.factor(res$method)
  res$method <- revalue(res$method, c("1" = "NaiveCFTP", "2" = "OptCFTP" , "3" = "AR"))
  return(res)
}

data_out <- generate.data()
res <- generate.data.frame(data_out)

ggplot(data = res[which(res$p == 0.01),], aes(x = factor(a), y = meanTosses, colour = factor(method))) + ggtitle("True p: 0.01") +
  geom_line(aes(group = factor(method))) + geom_point()
ggplot(data = res[which(res$p == 0.1),], aes(x = factor(a), y = meanTosses, colour = factor(method))) + ggtitle("True p: 0.1") +
  geom_line(aes(group = factor(method))) + geom_point()
ggplot(data = res[which(res$p == 0.25),], aes(x = factor(a), y = meanTosses, colour = factor(method))) + ggtitle("True p: 0.25") +
  geom_line(aes(group = factor(method))) + geom_point()
ggplot(data = res[which(res$p == 0.5),], aes(x = factor(a), y = meanTosses, colour = factor(method))) + ggtitle("True p: 0.5") +
  geom_line(aes(group = factor(method))) +  geom_point()
  #geom_ribbon(aes(ymin = minMeanTosses, ymax = maxMeanTosses, fill = factor(method)), alpha = 0.2, color = NA)

```

```{r old, echo=FALSE, eval=FALSE}
# for(a in a_seq) {
#   cat("Started a = ",a,"\n")
#   if(file.exists(paste0("efficiency_a_",a,".rds"))) {
#     #Recover previous work
#     cat("Recover previous work for a = ",a,"\n")
#     work_prev <- readRDS(paste0("efficiency_a_",a,".rds"))
#     res_prev <- work_prev$res
#     seed <- work_prev$seed
#     res[counter:(counter+nrow(res_prev)-1),] <- res_prev
#     counter <- counter + nrow(res_prev)
#     .Random.seed <- seed #Restore random seed
#   } else {
#     #Construct the ladder
#     cat("Constructing the ladder\n")
#     de <- DiceEnterprise$new(G=list(
#       list(1,matrix(c(a,0),nrow=1,ncol=2)),
#       list(1,matrix(c(0,a),nrow=1,ncol=2))
#     ))
#     #Get expected value for A-R
#     cat("Get expected tosses of AR\n")
#     exp_tosses_AR <- sapply(p_seq, function(p) {
#       C_p = p^a+(1-p)^a
#       Q <- 1
#       return(Q*a/C_p)
#     })
#     cat("Get empirical expected tosses of naive CFTP\n")
#     #Get value for naive CFTP
#     tosses_CFTP <- lapply(p_seq, function(p) {
#       return(de$sample(n=size_sample_CFTP, true_p = c(p,1-p), num_cores = num_cores, verbose = TRUE)[[2]])
#     })
#     exp_tosses_CFTP <- sapply(tosses_CFTP, mean)
#     cat("Get empirical expected tosses of augmented CFTP\n")
#     #Get an estimate of p
#     p_est_seq <- sapply(p_seq, function(p) {
#       mean(sample(c(1,0), prob = c(p,1-p), size = size_sample_est_p, replace = TRUE))
#     })
#     #Get a value for augmented CFTP
#     tosses_augmented_CFTP <- sapply(1:length(p_est_seq), function(i){
#       p_est <- p_est_seq[i]
#       p <- p_seq[i]
#       #Construct augmented ladder
#       new_de <- de$impose.efficiency(method="bound",true_p = c(p_est,1-p_est))
#       return(new_de$sample(n=size_sample_CFTP, true_p = c(p,1-p), num_cores = num_cores, verbose = TRUE)[[2]])
#     })
#     exp_tosses_augmented_CFTP <- sapply(tosses_augmented_CFTP, mean)
#     #Save in dataframe
#     initial_counter <- counter
#     for(i in 1:length(exp_tosses_CFTP)) {
#       res[counter,] <- c(p_seq[i],a,exp_tosses_CFTP[i],1)
#       res[counter+1,] <- c(p_seq[i],a,exp_tosses_augmented_CFTP[i],2)
#       res[counter+2,] <- c(p_seq[i],a,exp_tosses_AR[i],3)
#       counter <- counter + 3
#     }
#     
#     #Save to outer file
#     cat("Saving result for a = ",a,"\n")
#     res_temp <- res[initial_counter:(counter-1),]
#     saveRDS(list(res=res_temp,seed=.Random.seed), file = paste0("efficiency_a_",a,".rds"))
#   }
# }
# 
# res$method <- as.factor(res$method)
# res$method <- revalue(res$method, c("1" = "NaiveCFTP", "2" = "OptCFTP" , "3" = "AR"))
```

### 2019-01-15 inst/doc/DiceEnterprise.R

@@ -42,46 +42,24 @@ plot.confidence.interval(fp_tosses[[1]],print(bf$evaluate(3/4)))
 ## ----bf_error, error=TRUE------------------------------------------------
 bf_amp <- BernoulliFactory$new(f_1 = list(2,1), f_2 = list(c(1,-2),c(0,1)))
 
-## ----bf_logistic---------------------------------------------------------
-toss.coin.logistic <- function(n,p) {
-  sample(1:2, size = n, replace = TRUE, prob = c(p,1-p))
-}
-C <- 10
-p <- 0.01
-size_sample <- 10000
+## ----bf_threshold, error=TRUE--------------------------------------------
+bf_error <- BernoulliFactory$new(f_1 = list(c(-1,1,0.749),c(2,1,0)), 
+                                 f_2 = list(c(1,-1,0.251),c(2,1,0)))
+bf_no_error <- BernoulliFactory$new(f_1 = list(c(-1,1,0.749),c(2,1,0)), 
+                                 f_2 = list(c(1,-1,0.251),c(2,1,0)),
+                                 threshold = 500)
 
-bf_logistic <- BernoulliFactory$new(f_1 = list(C,1), f_2 = list(1,0))
-res <- bf_logistic$sample(n = size_sample, roll.fun = toss.coin.logistic, num_cores = 2, verbose = TRUE, p=p)
+## ----bf_AR---------------------------------------------------------------
+fp_tosses_AR <- bf$sample.AR(n = 1000, roll.fun = toss_coin)
+plot.confidence.interval(fp_tosses_AR,bf$evaluate(3/4))
 
-mean_Huber <- C/(1+C*p)
-print(mean_Huber) #Mean of Huber Algorithm
-print((1+C)/(1+C*p))  #Mean of Dice Enterprise
-print(mean(res[[2]]))
+## ----bf_AR_verbose1------------------------------------------------------
+fp_tosses_AR2 <- bf$sample.AR(n = 1000, roll.fun = toss_coin, verbose = TRUE)
+print(paste0("Empirical number of tosses required: ",mean(fp_tosses_AR2$empirical_tosses)))
 
-#CHECK IF IT IS CORRECT
-table(res[[1]])/size_sample
-print(c(C*p/(1+C*p), 1/(1+C*p)))
-
-
-
[…]

### 2019-01-15 inst/doc/Efficiency.R

@@ -1,0 +1,205 @@
+## ----setup, include = FALSE----------------------------------------------
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  fig.width=7, fig.height=4,
+  cache=TRUE
+)
+require("DiceEnterprise")
+library(plyr)
+
+## ----setup_efficiency, echo = FALSE--------------------------------------
+p_seq <- c(0.01,0.1,0.25,0.5) #Sequence of true values of p
+a_seq <- 1:10 #Sequence of a
+size_sample_CFTP <- 1000 #Size sample CFTP
+size_sample_est_p <- 1000 #Size tosses to get an estimate of p
+num_cores <- 2 #Number of cores used
+initial_seed <- 17 #Seed for RNG
+
+generate.data <- function() {
+  set.seed(initial_seed, "L'Ecuyer-CMRG") 
+  data_out <- vector("list", length = length(a_seq))
+  counter <- 1
+  for(a in a_seq) {
+    cat("Started a = ",a,"\n")
+    if(file.exists(paste0("efficiency_a_",a,".rds"))) {
+      #Recover previous work
+      cat("Recover previous work for a = ",a,"\n")
+      work_prev <- readRDS(paste0("efficiency_a_",a,".rds"))
+      data_out[[counter]] <- work_prev$data_out
+      counter <- counter+ 1
+      seed <- work_prev$seed
+      set.seed(seed, "L'Ecuyer-CMRG")  #Restore random seed
+    } else {
+      #Construct the ladder
+      cat("Constructing the ladder\n")
+      de <- DiceEnterprise$new(G=list(
+        list(1,matrix(c(a,0),nrow=1,ncol=2)),
+        list(1,matrix(c(0,a),nrow=1,ncol=2))
+      ))
+      #Get expected value for A-R
+      cat("Get expected tosses of AR\n")
+      exp_tosses_AR <- sapply(p_seq, function(p) {
+        C_p = p^a+(1-p)^a
+        Q <- 1
+        return(Q*a/C_p)
+      })
+      cat("Get empirical expected tosses of naive CFTP\n")
+      #Get value for naive CFTP
+      tosses_CFTP <- lapply(p_seq, function(p) {
+        return(1)
+        return(de$sample(n=size_sample_CFTP, true_p = c(p,1-p), num_cores = num_cores, verbose = TRUE)[[2]])
[…]

### 2019-01-15 inst/doc/Efficiency2.R

@@ -1,0 +1,164 @@
+## ----setup, include = FALSE----------------------------------------------
+knitr::opts_chunk$set(
+collapse = TRUE,
+comment = "#>",
+fig.width=7, fig.height=4,
+cache=TRUE
+)
+rm(list=ls())
+require("DiceEnterprise")
+library(plyr)
+
+## ----high_power_univariate_ladder----------------------------------------
+rm(list=ls())
+require("DiceEnterprise")
+
+p <- 1/2 #True value of p
+a <- 10 #Exponent
+n <- 1000 #Size sample CFTP
+num_cores <- 4 #Number of cores used
+initial_seed <- 17 #Seed for RNG
+increase_degree <- 260 #How much the degree is increased
+de_list <- vector("list", length = increase_degree+1) #Store all the de objects
+
+set.seed(initial_seed, "L'Ecuyer-CMRG") 
+#Construct the ladder
+cat("Constructing the dice enterprise \n")
+  de <- DiceEnterprise$new(G=list(
+    list(1,matrix(c(2*a,0),nrow=1,ncol=2)),
+    list(1,matrix(c(a,a),nrow=1,ncol=2)),
+    list(1,matrix(c(0,2*a),nrow=1,ncol=2))
+  ))
+  de_list[[1]] <- de
+  cat("Constructing the augmented dice enterprises \n")
+  for(i in 1:increase_degree) {
+    #Check if it already exists
+    if(file.exists(paste0("efficiency_uni_data/ladder_a_",a,"_degree_",i,".rds"))) {
+      cat("Restoring previous work for degree +",i," \n")
+      prev_work <- readRDS(paste0("efficiency_uni_data/ladder_a_",a,"_degree_",i,".rds"))
+      de_list[[i+1]]  <- prev_work$de_aug
+    } else {
+      cat("Constructing the augmented dice enterprises for +",i," \n")
+      if(i == 1) {
+        de_list[[i+1]] <- de$increase.degree(1)
+      } else {
+        de_list[[i+1]] <- de_list[[i-1]]$increase.degree(1)
+      }
+      #Save object
+      saveRDS(list(de_aug=de_list[[i+1]],i=i), file = paste0("efficiency_uni_data/ladder_a_",a,"_degree_",i,".rds"))
+    }
+  
+}
+
[…]

### 2019-01-15 inst/doc/PaperExamples.R

@@ -11,3 +11,300 @@ require("ggplot2")
 require("xtable")
 set.seed(10,"L'Ecuyer-CMRG")
 
+## ----bf_settings---------------------------------------------------------
+sample_size <- 1000 #Number of tosses
+probs_vec <- c(0.01,0.1,0.25,0.5,0.75,0.9,0.99) #Vector of true probabilities of the p-coin
+num_cores <- 2 #Number of cores used (parallel computing is not supported on Windows)
+conf_level <- 0.95 #Confidence interval level
+
+## ----bf_toy, echo=FALSE, results='asis'----------------------------------
+toss.coin <- function(n, p) {
+  sample(1:2, size = n, replace = TRUE, prob = c(p,1-p))
+}
+bf <- BernoulliFactory$new(f_1 = list(coeff = c(sqrt(2)), power = c(3)), 
+                           f_2 = list(coeff = c(-5,11,-9,3), power = c(3,2,1,0))) 
+
+bf.res <- function(toss.coin, bf) {
+  bf_res <- matrix(NA, nrow = 6, ncol = length(probs_vec))
+  rownames(bf_res) <- c("f(p)", "LowerCI", "Emp.f(p)", "UppCI", "Exp.TossesNoDoubling", "Exp.TossesDoubling")
+  colnames(bf_res) <- as.character(probs_vec)
+  for(i in 1:length(probs_vec)) {
+    sample_res <- bf$sample(n = sample_size, roll.fun = toss.coin, p = probs_vec[i], num_cores = num_cores, verbose = TRUE, double_time = FALSE)
+    sample_res_double_time <- bf$sample(n = sample_size, roll.fun = toss.coin, p = probs_vec[i], num_cores = num_cores, verbose = TRUE, double_time = TRUE)
+    #Get lower and upper CI
+    if(isTRUE(all.equal(sample_res[[1]],rep(1,sample_size)))) {
+      #All equal to 1 (Heads)
+      lower_ci <- 1
+      upper_ci <- 1
+    } else if(isTRUE(all.equal(sample_res[[1]],rep(2,sample_size)))) {
+      #All equal to 2 (Tails)
+      lower_ci <- 0
+      upper_ci <- 0
+    } else {
+      ci <- MultinomCI(table(sample_res[[1]]),conf.level = conf_level)
+      lower_ci <- ci[1,2]
+      upper_ci <- ci[1,3]
+    }
[…]
